\textsc{RILS}-\textsc{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-11-27 paper_resources/doc/paper.tex
Paper polishing.

For the (13) competitors, the results are exported from SRBench \url{https://cavalab.org/srbench/results/}, reported in~\cite{la2021contemporary}. 
The maximum computation time allowed for each run of \textsc{RILS}-\textsc{ROLS} is set to 1 hour, while the maximal number of fitness function evaluations is set to 1 million. All other algorithms from SRBench used the time limit that was equal or bellow 8 hours runtime and the same fitness evaluation limit of 1 million (see~\cite{la2021contemporary}). We did not use the full time limit of 8 hours because it was not necessary -- we empirically concluded that there almost no improvements after 1 hour \textsc{RILS}-\textsc{ROLS} runtime.

@@

SRBench is an open-source benchmarking project which merge a large set of diverse benchmark datasets, contemporary SR methods as well as ML methods around a shared model evaluation and an environment for analysis, see~\cite{la2021contemporary}. It brings researchers an easy handled environment to evaluate, compare, and analyze their methods.  We involve two real-world inspired ground-truth benchmark sets from SRBench, that represent various types of regression problems, into our computational evaluation: 
\begin{itemize}
	\item \textsc{Feynman} benchmark set, inspired from physics and formulas/models that describes   various natural laws.  
	It consists of $10^5$  samples, each described in~\cite{udrescu2020ai}. Some exact models (equations) of problem instances from this benchmark are listed in Table~\ref{tab:Feynamn-Eq}.  Among them, 116 regression problem instances are chosen in the experimental evaluation, with accordance to the results reported in~\cite{la2021contemporary} that used the same problem instances.

@@

	\item \textsc{Strogatz} benchmark set from~\cite{la2016inference}. 
	Each dataset is represented by a 2-state system of first-order, ordinary differential equations. 
	The aim of each problem is to predict rate of change of the subsequent state w.r.t\  in dependence of the current two states. These equations describes natural processes modelled by non-linear dynamics  exhibiting chaos.  The equations for some of the datasets that belong to this benchmark are shown in Table~\ref{table:strogatz-ODEs}. These problem instances are downloadable from \url{https://github.com/lacava/ode-strogatz}. In overall, 14 ground-truth models are used to generate the problem instances used in the comparison, also in accordance with~\cite{la2021contemporary}.

@@

\end{itemize}
The above-mentioned benchmark sets may be biased since the models which describe physical laws usually impose   various symmetries, periodicity w.r.t. some variables, internal separability on some variables etc. In that way, the search process and the search space can be pre-defined and thus significantly reduced. Thereby, some methods may get significant benefit with this regard, while some others may be in a subordinate situation. 
In order to permit any kind of favoritism caused by physical lows behind the input data, we generate a set of randomly generated (non-biased) SR problem instances; this benchmark set is therefore labelled by \textsc{Random}. These instances are generated according the following procedure. \fxnote{Opisati proceduru generisanja instanci + kombinaciju parametara za instance, itd.}

@@

\section{Overview of \textsc{RILS}-\textsc{ROLS}  python package}\label{sec:appendix-1}
\fxnote{TODO}

>>>>

For the (13) competitors, the results are exported from SRBench \url{https://cavalab.org/srbench/results/}, reported in~\cite{la2021contemporary}. 
The maximum computation time allowed for each run of \textsc{RILS}-\textsc{ROLS} is set to 1 hour, while the maximal number of fitness function evaluations is set to 1 million. All other algorithms from SRBench used the time limit that was equal or bellow 8 hours runtime and the same fitness evaluation limit of 1 million (see~\cite{la2021contemporary}). We did not use the full time limit of 8 hours because we empirically concluded that there were almost no improvements after 1 hour \textsc{RILS}-\textsc{ROLS} runtime.

@@

SRBench (\url{https://cavalab.org/srbench/}) is an open-source benchmarking project which merges a large set of diverse benchmark datasets, contemporary SR methods as well as ML methods around a shared model evaluation and an environment for analysis, see~\cite{la2021contemporary}. SRBench considers two types of symbolic regression problems: 1) \emph{ground-truth} problems, for which the exact model is known, and 2) \emph{black-box} problems, for which the exact model is not known. In our work, we consider only the former one, since \textsc{RILS-ROLS} was not designed to solve \emph{black-box} problems. There are two groups of instances in SRBench \emph{ground-truth} problem set:
\begin{itemize}
	\item \textsc{Feynman} instances are inspired by physics and formulas/models that describes various natural laws.  
	There are 116 instances, where each one consists of $10^5$  samples (see \cite{udrescu2020ai} for more details). Some exact models (equations) of \textsc{Feynman} instances are listed in Table~\ref{tab:Feynamn-Eq}.

@@

	\item \textsc{Strogatz} instances are introduced in \cite{la2016inference}. 
	Each instance represents a 2-state system of first-order, ordinary differential equations. 
	The aim of each problem is to predict the rate of change of the subsequent state. These equations describe natural processes modeled by non-linear dynamics exhibiting chaos.  The equations for some of the datasets that belong \textsc{Strogatz} are shown in Table~\ref{table:strogatz-ODEs}. In overall, there are 14 \textsc{Strogatz} instances.

@@

\fxnote{ACA: do ovde sam pregledao}
The above-mentioned benchmark sets may be biased since the models which describe physical laws usually impose   various symmetries, periodicity w.r.t. some variables, internal separability on some variables etc. In that way, the search process and the search space can be pre-defined and thus significantly reduced. Thereby, some methods may get significant benefit with this regard, while some others may be in a subordinate situation. 
In order to permit any kind of favoritism caused by physical lows behind the input data, we generate a set of randomly generated (non-biased) SR problem instances; this benchmark set is therefore labelled by \textsc{Random}. These instances are generated according the following procedure. \fxnote{Opisati proceduru generisanja instanci + kombinaciju parametara za instance, itd.}

@@

\section{Overview of \textsc{RILS}-\textsc{ROLS} python package}\label{sec:appendix-1}

\textsc{RILS-ROLS} algorithm is available for install on the well-known Python package repository \url{https://pypi.org}, so it can be easily installed by typing commands:
\begin{verbatim}
	pip install rils-rols
\end{verbatim}
\textsc{RILS-ROLS} project page is at \url{https://pypi.org/project/rils-rols}. Here, one can find a minimal working example on how-to-use \textsc{RILS-ROLS}:

\begin{verbatim}
from rils_rols.rils_rols import RILSROLSRegressor
from math import sin, log

regr = RILSROLSRegressor()

# toy dataset 
X = [[3, 4], [1, 2], [-10, 20], [10, 10], [100, 100], [22, 23]]
y = [sin(x1)+2.3*log(x2) for x1, x2 in X]

# RILSROLSRegressor inherits BaseEstimator (sklearn)
regr.fit(X, y)

# this prints out the learned model
print("Final model is "+str(regr.model))

# applies the model to a list of input vectors
X_test = [[4, 4], [3, 3]]
y_test = regr.predict(X_test)
print(y_test) 
\end{verbatim}

Python sources, experimental results and all other \textsc{RILS-ROLS} resources can be found at the project GitHub page \url{https://github.com/kartelj/rils-rols}.
